Data preparation

### 2023-03-15 04. Analysis/scripts/01_data_preparation.Rmd
Update 01_data_preparation.Rmd
Added headers, include section comparing combined dataframe to originals, began to address nulls and blanks.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
```

@@

### Join data

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
```

@@

## Joining data

### Join 1

@@

### Join 2

@@

## Compare to original dataframes

```{r}
prep_base_1_summary <- 
  prep_base_1 %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(employee_name),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_base_1")
```

```{r}
prep_base_2_summary <- 
  prep_base_2 %>% 
  # unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(full_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address)
  ) %>% 
  mutate(dataframe = "prep_base_2")
```

```{r}
prep_building_summary <- 
  prep_building %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>%
  summarise(
    rows = n(),
    person_id = n_distinct(person_id),
    employee_name = n_distinct(employee_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_building")
```

```{r}
prep_summary <- 
  prep_base_1_summary %>% 
  bind_rows(prep_base_2_summary) %>% 
  bind_rows(prep_building_summary) %>% 
  pivot_longer(-dataframe) %>% 
  pivot_wider(names_from = dataframe,values_from = value)
```

```{r}
prep_summary %>% 
  flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

```{r}
prep_base_1_rows <- prep_summary %>% filter(name == 'rows') %>% pull(prep_base_1)

prep_base_1_employee_name <- prep_summary %>% filter(name == 'employee_name') %>% pull(prep_base_1)

diff_rows_count <- 
  (prep_base_1_rows) -
  (prep_summary %>% filter(name == 'rows') %>% pull(prep_building))

diff_rows_prop <- 
  diff_rows_count/prep_base_1_rows

diff_employee_name_count <- 
  (prep_base_1_employee_name) -
  (prep_summary %>% filter(name == 'employee_name') %>% pull(prep_building))

diff_employee_name_prop <- 
  diff_employee_name_count/prep_base_1_employee_name
```

From this table, we find:

1.    A loss of `r scales::comma(diff_rows_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_rows_prop)`
2.    A loss of `r scales::comma(diff_employee_name_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_employee_name_prop)`

These losses of data are reasonable.

Create a unique dataframe of just `date` and `person_id`.

```{r}
prep_uniq <- 
  prep_building %>% 
  select(date,person_id)
```

## Addressing blanks, nulls, and other values

Created a dataframe of variable metadata.

```{r}
column_meta_data <- 
  prep_building %>% 
  contents()

column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = case_when(NAs > 0 ~ TRUE,
                                  NAs == 0 ~ FALSE)
  )

Counting top 10 values for all character or factor columns to identify other common values.

```{r}
values_count <- 
  prep_building %>% 
  select(where(is.character) | where(is.factor)) %>% 
  names()

values_count %>% 
  map(~count(prep_building,!!sym(.)) %>% 
  arrange(desc(n)) %>%
  slice_head(n = 10))

```

No obvious values.

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

Checking blanks for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

Create dataframe of only variables that have blanks.

```{r}
problem_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable | nulls_by_variable)

problem_variables
```

```{r}
problem_variables <- 
  problem_variables %>% 
  mutate(group = 
           case_when(variable == "position_category_code" ~ "position_category_code",
                     variable == "position_category_description" ~ "position_category_code",
                     TRUE ~ variable
                     ))

problem_variables
```

```{r}
problem_groups <- 
  problem_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

problem_groups
```
Also create a building dataframe to add other dataframes where missing have been removed.

```{r}
prep_problem <- prep_uniq
```

### Problems removal

#### Step `r step_n`: `business_line_2_letter_department_id`

Removing blanks produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_point() +
  geom_smooth()
```

Increase in July.

Is this an important variable?

No. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

#### Step `r step_n`: `employee_e_mail_address`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random.

Is this an important variable?

Yes, but we won't be using it directly in the analysis. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

#### Step `r step_n`: `location_city`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Weird that early in the year is low, then gets steady later.

Is this an important variable?

Yes, but we have other location variables to rely on. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

#### Step `r step_n`: `location_state`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random, and low numbers.

Is this an important variable?

Yes, but we have other location variables to rely on. It appears from above that these are international locations. We'll replace nulls with "(International)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(International)",
                        TRUE ~ .)
    )
  )

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

# CONTINUE FROM HERE.
